Fast Bayesian Inference for Exoplanet Discovery in Radial Velocity Data

### 2014-10-26 Paper/ms.tex
Keywords
Forceful words

\begin{keywords}
methods: data analysis
\end{keywords}

@@

The number of orbiting planets, $N$, is an important parameter.
To calculate the posterior distribution for $N$, most authors
consider various trial values of $N$, and calculate the marginal likelihood
\begin{eqnarray}
p(\bdata | N) &=& \int p(\btheta | N) p(\bdata | \btheta, N) \, d^n \theta
\end{eqnarray}
for each possible value of $N$
\citep[e.g.][]{2011MNRAS.415.2523G, 2014MNRAS.437.3540F, fengji}.
The posterior distribution for $N$ can then be found straightforwardly by
using Bayes' rule with $N$ as the only unknown parameter:
\begin{eqnarray}
p(N | \bdata) &=& \frac{p(N)p(\bdata | N)}{\sum_N p(N)p(\bdata | N)}.
\end{eqnarray}
Popular
methods for calculating the marginal likelihood are Nested Sampling
\citep{skilling} and ideas related to thermodynamic integration
\citep[e.g.][]{neal}.

>>>>

\begin{keywords}
stars: planetary systems --- techniques: radial velocities ---
methods: data analysis --- methods:statistical
\end{keywords}

@@

\subsection{Inferring the number of planets}
The number of orbiting planets, $N$, is an important parameter.
To calculate the posterior distribution for $N$, most authors
consider various trial values of $N$, and calculate the marginal likelihood
\begin{eqnarray}
p(\bdata | N) &=& \int p(\btheta | N) p(\bdata | \btheta, N) \, d^n \btheta
\end{eqnarray}
for each possible value of $N$
\citep[e.g.][]{2011MNRAS.415.2523G, 2014MNRAS.437.3540F, fengji}, marginalising
over all other model parameters.
The posterior distribution for $N$ can then be found straightforwardly by
using Bayes' rule with $N$ as the only unknown parameter:
\begin{eqnarray}
p(N | \bdata) &=& \frac{p(N)p(\bdata | N)}{\sum_N p(N)p(\bdata | N)}.
\end{eqnarray}
Popular
methods for calculating the marginal likelihood are Nested Sampling
\citep{skilling} and ideas related to thermodynamic integration
\citep[e.g.][]{neal}.

This traditional approach can be very time consuming.
Methods for calculating the marginal likelihood are
already more intensive than standard MCMC methods for sampling the posterior,
because they usually involve a sequence of probability distributions
(e.g. the constrained priors in Nested Sampling, or the annealed distributions
in thermodynamic integration) rather than a single distribution (the posterior).
This intensive process needs to be run many times, for $N=0$, $N=1$, $N=2$, and
so on.

The traditional approach also contradicts standard practice in Bayesian
computation. Imagine we are trying to compute the posterior distribution for
a parameter $a$ in the presence of a nuisance parameter $b$. This is solved
by exploring the joint posterior for $a$ and $b$, and then only looking at the
generated values of $a$. Nobody would suggest the wasteful alternative
of using a discrete grid of possible $a$ values and doing an entire Nested
Sampling run for each, to get the marginal likelihood as a function of $a$.
Yet when the hypothesis space for $a$ is already discrete, the wasteful
alternative somehow becomes standard practice.
